Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques

### 2018-02-07 paper/gloaguenNmfTraffic.tex
updates in font size of figures

\author{
    Jean-Rémy GLOAGUEN\\
    Arnaud Can\\
    LAE\\
    Ifsttar\\
    jean-remy.gloaguen@ifsttar.fr
  \and
    Mathieu Lagrange\\
	Jean-François Petiot \\
    LS2N, CNRS\\
    \'Ecole Centrale de  Nantes\\
}
\date{}
\begin{document}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf_2.pdf}
\caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c))}
\label{fig:example_NMF}
\end{figure}

@@

with the \textit{Traffic Interference Ratio} $TIR \in$ [-12 -6 0 6  12] dB. When $TIR < 0$ dB, the traffic component is less present than the interfering class. In the opposite, for $TIR > 0$ dB, the traffic class is louder than the interfering class. The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values). The corpora are available for download at: \url{https://zenodo.org/record/1145855#.Wl2oPnkiGos}

@@

The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ([-12 -6 0 6 12] dB). The spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap, see Figure \ref{fig:bloc_experiment}.

@@

Assuming that the traffic spectral profile is largely concentrated in the low frequency components, a first estimator to determine the traffic sound level is a frequency low-pass filter. It depends only on the cut-off frequencies $f_c \in $ [500 1k 2k 5k 10k 20k] Hz. The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq. \ref{eq:v_tr_filtered}),

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\hline
    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modality}\end{tabular}\\ \toprule
\end{tabularx}

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}@{}C{2cm}}
	\hline
    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} &\begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modality}\end{tabular}\\ \toprule
\end{tabularx}

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
   $\mathbf{t}$  &  from 0.30 to 0.70 with 0.01 step & 40\\
   \bottomrule
\end{tabularx}
\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator}
\label{tab:experimental_factorsNMF}
\end{table*}

@@

In order to prevent potential overfitting issues, the dictionary is built from a separate sound database dedicated specifically to this task. It is composed of 53 audio files of passing cars. These recordings have been made on the Ifsttar's runway too, with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio). First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$). Then time/frequency windows of $F \times w_t $ dimensions are applied without overlapping on the spectrogram in order to consider several spectra for each audio file where $w_t \in [0.5~1]$ second. In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectrum a of $F \times 1$ dimension. With this size of window, it is possible to obtain the characteristic pitches of the different audio samples. One obtain for each value of $w_t$, from the 53 audio samples of passing cars respectively 2218 and 1109 elements. Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering algorithm is applied to reduce the number of spectra to $K \in \left[ 25,~50,~100,~200\right]$. The $K$ centroids are then the elements considered in the dictionary. A special case is added where the root mean square of \textit{all} the spectrogram is applied ($w_t$ = $all$) in order to build a dictionary with the spectral envelope of each audio sample. In this case, 53 spectra are obtained. The $K$-means clustering algorithm is then reduced to $K \in \left[25~50 \right]$.

@@

Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed. For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ [0.5 1] s whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning}. For TI-NMF, 24000 combinations are computed. In all, 25380 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.

@@

First, the errors produced by the filter are detailed. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources. Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained. It is then considered as the baseline to compare the performances of NMF.

@@

    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the M scenes of the sub-classes \textit{alert} at $TIR = -12$ dB(\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12})}
    \label{fig:dictionaryExtraction}
\end{figure*}

Compared with the filter errors, the choice of some NMF approaches makes it possible to decrease the error of the road traffic sound level estimation. The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline. Sem-NMF and TI-NMF have better results. The lowest average error is obtained for TI-NMF, 2.19 $\pm$ 2.18, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms. On the other hand, the semi-supervised approach has a higher error but a lower standard deviation (2.32 $\pm$ 1.26).

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2}
                \label{fig:TIR_class_sup}
    \end{subfigure}

@@

Finally, TI-NMF with a threshold $t = 0.42$ and $\beta$ = 1, offers the lowest average error (Table \ref{tab:results}). Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments. The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal. In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ [-12, 12] dB.

@@

For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decreases the error (< 1 dB). The kept elements are then more suited to the scenes than a fixed dictionary. The error for these $TIR$ is then due to the thresholding which put aside some elements that are nevertheless related to the traffic component. With a low threshold, it is possible to decrease the error. For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30. In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease then to 4.46 $\pm$ 1.66. In order to generalize this method where no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 and is the one that best balanced these opposite cases.

>>>>

\author{
    Jean-Rémy Gloaguen\\
    Arnaud Can\\
    LAE\\
    Ifsttar\\
    jean-remy.gloaguen@ifsttar.fr
  \and
    Mathieu Lagrange\\
	Jean-François Petiot \\
    LS2N, CNRS\\
    \'Ecole Centrale de  Nantes\\
}
\date{}
\begin{document}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
\caption{Example of a simple NMF for urban sound mixture composed of 3 sound events (car passages, car horn and bird's whistles), $\mathbf{W}$ is composed of 3 elements too ($K$ = 3) which correspond in 3 audio spectra (car passages (a), car horn (b) and bird's whistles (c))}
\label{fig:example_NMF}
\end{figure}

@@

with the \textit{Traffic Interference Ratio} $TIR \in$ $\lbrace$12, -6, 0, 6, 12$\rbrace$ dB. When $TIR < 0$ dB, the traffic component is less present than the interfering class. In the opposite, for $TIR > 0$ dB, the traffic class is louder than the interfering class. The total number of scenes designed is 750 (6 sub-corpus $\times$ 25 scenes $\times$  5 $TIR$ values). The corpora are available for download at: \url{https://zenodo.org/record/1145855#.Wl2oPnkiGos}

@@

The experiment consists in estimating the road traffic sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) composed each of 25 audio files ($M$ = 25) and for 5 $TIR$ ($\lbrace$-12, -6, 0, 6, 12$\rbrace$ dB). The spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap, see Figure \ref{fig:bloc_experiment}.

@@

Assuming that the traffic spectral profile is largely concentrated in the low frequency components, a first estimator to determine the traffic sound level is a frequency low-pass filter. It depends only on the cut-off frequencies $f_c \in  \lbrace$500, 1k, 2k, 5k, 10k, 20k$\rbrace$ Hz. The spectrogram $\mathbf{V}$ is filtered and the remaining energy is then considered as traffic component (eq. \ref{eq:v_tr_filtered}),

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
	\hline
    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} & \begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modalities}\end{tabular}\\ \toprule
\end{tabularx}

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}@{}C{2cm}}
	\hline
    \textbf{\begin{tabular}[c]{@{}l@{}}experimental \\ factors\end{tabular}} & \textbf{modalities} &\begin{tabular}[c]{@{}C{2cm}@{}}\textbf{number}\\ \textbf{of modalities}\end{tabular}\\ \toprule
\end{tabularx}

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{12cm}@{}C{2cm}@{}}
   threshold $\mathbf{t}$  &  from 0.30 to 0.70 with 0.01 step & 41\\
   \bottomrule
\end{tabularx}
\caption{Summary of the different experimental factors and their modalities taken into account in NMF estimator}
\label{tab:experimental_factorsNMF}
\end{table*}

@@

In order to prevent potential overfitting issues, the dictionary is built from a separate sound database dedicated specifically to this task. It is composed of 53 audio files of passing cars. These recordings have been made on the Ifsttar's runway too, with the same experimental conditions that the recordings of the \textit{SimScene} database but with two different cars (Dacia Sandero and Renault Clio). First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$). Then time/frequency windows of $F \times w_t $ dimensions are applied without overlapping on the spectrogram in order to consider several spectra for each audio file where $w_t \in \lbrace$0.5, 1$\rbrace$ second. In each window, the root mean square value is calculated on each frequency bin to reduce the windowed spectrogram in one spectrum a of $F \times 1$ dimension. With this size of window, it is possible to obtain the characteristic pitches of the different audio samples. One obtains for each value of $w_t$, from the 53 audio samples of passing cars respectively 2218 and 1109 elements. Since the number of elements given by this processing is high, in order to reduce the computational time and avoid redundant information, a $K$-means clustering algorithm is applied to reduce the number of spectra to $K \in \lbrace$25, 50, 100, 200$\rbrace$. The $K$ centroids are then the elements considered in the dictionary. A special case is added where the root mean square of \textit{all} the spectrogram is applied ($w_t$ = $all$) in order to build a dictionary with the spectral envelope of each audio sample. In this case, 53 spectra are obtained. The $K$-means clustering algorithm is then reduced to $K \in \lbrace$25, 50$\rbrace$.

@@

Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factorsFilter}  between the 5 levels of $TIR$, the 6 sub-classes and the 6 cut-off frequencies $f_c$, 180 settings are performed (6 $\times$ 5 $\times$ 6). For Sup-NMF and Sem-NMF, according to Table \ref{tab:experimental_factorsNMF},  1200 associations of factors are made where the 4 levels of $K$ are associated with $w_t \in$ $\lbrace$0.5 1$\rbrace$ second whereas only 2 levels of $K$ (25 and 50) are associated with $w_t = all$, see part \ref{part:dictionary_learning} (6 $\times$ 5 $\times$ 2 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 2). For TI-NMF, 24600 combinations are computed (6 $\times$ 5 $\times$ (2 $\times$ 4 + 1 $\times$ 2) $\times$ 41). In all, 25980 settings are performed between the different forms of the dictionary $\mathbf{W}$ and the multiple experimental factors.

@@

First, the errors produced by the filter are detailed. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources. Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lowest mean error obtained (2.89 ($\pm$ 2.84) dB). It is then considered as the baseline to compare the performances of NMF.

@@

    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the M scenes of the sub-classes \textit{alert} at $TIR = -12$ dB (\subref{fig:alert_-12}) and at $TIR$ = 12 dB (\subref{fig:alert_12})}
    \label{fig:dictionaryExtraction}
\end{figure*}

Compared with the filter errors, the choice of some NMF approaches makes it possible to decrease the error of the road traffic sound level estimation. The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline. Sem-NMF and TI-NMF have better results. The lowest average error is obtained for TI-NMF, 2.19 ($\pm$ 2.18) dB, for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms. On the other hand, the semi-supervised approach has a higher error but a lower standard deviation (2.32 ($\pm$ 1.26) dB).

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
        \caption{$MAE$ error for each $TIR$ and sub-class with the frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
        \caption{$MAE$ error for each $TIR$ and sub-class with Sup-NMF and $\beta$ = 2}
                \label{fig:TIR_class_sup}
    \end{subfigure}

@@

Finally, TI-NMF with a threshold $t = 0.42$ and $\beta$ = 1, offers the lowest average error (Table \ref{tab:results}). Unlike Sup-NMF, where $\mathbf{W}$ is fixed, and Sem-NMF, where only $\mathbf{W_r}$ is updated, TI-NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments. The closest elements of the traffic component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal. In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR \in $ $\lbrace$-12, 12$\rbrace$ dB.

@@

For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decreases the error (< 1 dB). The kept elements are then more suited to the scenes than a fixed dictionary. The error for these $TIR$ is then due to the thresholding which put aside some elements that are nevertheless related to the traffic component. With a low threshold, it is possible to decrease the error. For $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30. In opposite for $TIR$ = -12 dB, it is better to choose a high threshold $t$ = 0.55, the error decrease then to 4.46 ($\pm$ 1.66) dB. In order to generalize this method where no prior knowledge on the urban environment, the chosen threshold $t$ is then fixed to $t$ = 0.42 as it is the one that best balanced these opposite cases.
